---
title: "RandomForest approch with `caret` package and parallel training"
author: "onion"
date: "20170401"
output: html_document
---

> ## Introduction
> * Data cleaning
> * Data exploration
> * Data slicing
> * Model fitting
> * Making submission 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,message=FALSE,warning=F)
```

## 1.Data cleaning: loading data and imputing

First load the data and implement packages. Here I use `data.table` to manipulate the data. Combine train set and test set as a single full set.
```{r}
library(caret)
library(data.table)
library(ggplot2)
library(doParallel)
data=rbind(fread("data/train.csv"),fread("data/test.csv"),fill=T)
```


Then let's find out the missing data. We calculate the number of `NA` in each column and also, noting that some character variables with some records equal to "", we should take them into consideration too.

```{r}
colSums(sapply(data,is.na))
sapply(data,FUN = function(x) length(which(x=="")))
```

The "Embarked" is related to "Pclass" and "Fare", therefore the missing "Embarked" might be "C" according to the boxplot of "Fare" by Pclass groups. 

```{r}
ggplot(data,aes(x=Pclass,y=Fare,group=Pclass,fill=as.factor(Pclass)))+geom_boxplot()
data[Embarked==""]
data[Embarked=="",Embarked:="C"]
```

The missing "Fare" can be imputed by the mean of records with same "Embarked" and "Pclass". 

```{r}
data[is.na(Fare),Fare:=data[Pclass==data[is.na(Fare),Pclass]&&Embarked==data[is.na(Fare),Embarked],mean(Fare,na.rm=T)]]
```

As for "Age", the information hide within the title in "Name", such as "Mr.", "Mrs.". We infer "Age" by the mean of people with same title.

```{r}
getTitle=function(s){
  reg=sapply(s,FUN = function(x) return(gregexpr("\\s\\S+\\.",x)))
  return(sapply(1:length(s),FUN=function(x) substr(s[x],reg[x][[1]]+1,reg[x][[1]]+attr(reg[x][[1]],"match.length")-1)))
}
data[,`:=`(Title=getTitle(Name))]
data[,gpAge:=mean(Age,na.rm=T),by=Title]
data[is.na(Age),Age:=gpAge,]
```

## 2.Data exploration: feature engineering

There exists a surviving penalty for adults according to the density plot. We create new variable "Child" which contains 0 for adults and 1 for children. Sum up "SibSp" and "Parch" to create a variable describing the family size. Then transform categories variables to factors.

```{r}
ggplot(data=data[!is.na(Survived)],aes(x=Age,colour=as.factor(Survived)))+geom_density()
table(data[(!is.na(Survived))][Age<18,Survived])
table(data[(!is.na(Survived))][Age>=18,Survived])
data[,`:=`(FamilySize=SibSp+Parch+1,Child=rep(0,.N))]
data[Age<18,Child:=1]
data[,c("Survived","Pclass","Sex","Embarked","Child")]=lapply(data[,c("Survived","Pclass","Sex","Embarked","Title","Child")], function(x) as.factor(x))
```

## 3.Data slicing

In order to choose the best model, we divide the train set into two sets: `training` for training the model and `testing` for validating. The proportion of records in `training` is 75%.

```{r}
train=data[!is.na(Survived)]
test=data[is.na(Survived)]
set.seed(1)
inTrain=createDataPartition(y=train[,Survived],p=0.75,list=FALSE)
training=train[inTrain,]
testing=train[-inTrain,]
```

## 4.Model fitting

We choose randomForest as the model. For faster training, we use `doParallel` package to accelerate the computing by multicores.

```{r}
cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)
fitControl <- trainControl(allowParallel = TRUE)
set.seed(1)
modelFit=train(Survived~Pclass+Sex+Age+SibSp+Parch+Fare+Embarked+FamilySize+Child,
               method="rf",data=training,trControl = fitControl)
stopCluster(cluster)
registerDoSEQ()
modelFit
```

How about the performance in `testing`?

```{r}
confusionMatrix(testing$Survived,predict(modelFit,testing))
```

## 5.Making submission

```{r}
cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)
fitControl <- trainControl(allowParallel = TRUE)
set.seed(1)
modelFit=train(Survived~Pclass+Sex+Age+SibSp+Parch+Fare+Embarked+FamilySize+Child,
               method="rf",data=train,trControl = fitControl)
stopCluster(cluster)
registerDoSEQ()
# write submission file
submit=test[,.(PassengerId)]
submit[,`:=`(Survived=predict(modelFit,test))]
```


